### 2010-08-12 doc/tech_doc/gridftpd/gridftpd.tex
assign document number
add description of log and pid options and spellcheck

{\raggedleft NORDUGRID-TECH-XX\par}

@@

With the evolution of Grid technologies, and the need for more
standardized interfaces to allow different Grid products to
communicate with each other, the Hosting Environment Daemon
(HED)~\cite{HED} framework and associated Advanced Resource Execution
Service (A-REX) web service were developed to replace the aging GFS
and GM with a modern standards-compliant product. However, the
transistion to a new framework takes time and so the legacy GFS is
still supported and maintained until such time as it becomes
obsolete. This document is designed for administrators who wish to run
the GFS alongside A-REX and describes how to configure and run the GFS
service. For installation of the GFS, up to date instructions may be
found on the NorduGrid website \textbf{\url{http://www.nordugrid.org}}.

@@

To make GFS to interoperate with other parts of ARC only one
\emph{jobplugin.so} needs to be configured.

@@

\begin{itemize}
\item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port
  number.  Default is 2811.
\item \textbf{\textit{include}}\textit{=path} - include contents of
  another file. Generic commands cannot be specified there.
\item \textbf{\textit{encryption}}\textit{=yes|no} - specifies if
  server will allow data transfer to be encrypted. Default is yes.
\item \textbf{\textit{pluginpath}}\textit{=path} - specifies the path
  where plugin libraries are installed. In a normal installation this
  is \emph{\$(ARC\_LOCATION)/lib/arc}.
\item \textbf{\textit{allowunknown}}\textit{=yes|no} - if set to
  \emph{yes}, clients are not checked against the grid-mapfile. Hence
  only access rules specified in this configuration file will be
  applied.
\item \textbf{\textit{firewall}}\textit{=hostname} - use IP address of
  the \textit{hostname} in response to PASV command instead of IP
  address of a network interface of the computer. An IP address can be
  used instead of \textit{hostname}. This command may be useful if the
  server is situated behind a NAT.
\item \textbf{\textit{unixgroup}}\textit{=group rule} - define local
  UNIX user and optionally UNIX group to which user belonging to
  specified authorization \textit{group} is mapped (see
  Section~\ref{sec:Authorization} for definition of group). Local
  names are obtained from the specified \textit{rule}. If the
  specified rule could not produce any mapping, the next command is
  used. Mapping stops at first matched rule. The following rules are
  supported:

@@

\begin{itemize}
\item \%D - subject of users's cerificate,
\item \%P - name of credentials' proxy file.
\end{itemize}

@@

\begin{itemize}
\item {[}\textbf{\textit{subject}}{]}\textit{=subject}
  {[}\textit{subject} {[}...{]}{]} - match user with one of specified
  subjects
\item \textbf{\textit{file}}={[}\textit{filename} {[}...{]}{]} - read
  rules from specified files (format of file is similar to Globus
  grid-mapfile with user names ignored)
\item \textbf{\textit{remote}}={[}\textit{ldap://host:port/dn}
  {[}...{]}{]} - match user listed in one of specified LDAP
  directories (uses network connection hence can take time to process)
\item \textbf{\textit{voms}}\textit{=vo group role capabilities} -
  accept user with VOMS proxy with specified \textit{vo},
  \textit{group}, \textit{role} and \textit{capabilities}. '{*}' can
  be used to accept any value
\item \textbf{\textit{vo}}={[}\textit{vo} {[}...{]}{]} - match user
  belonging to one of specified Virtual Organizations as defined in
  \emph{vo} configuration section (see below).
\item \textbf{\textit{group}}={[}\textit{groupname}
  {[}\textit{groupname} {[}...{]}{]}{]} - match user already belonging
  to one of specified groups.
\item \textbf{\textit{plugin}}\textit{=timeout plugin}
  {[}\textit{arg1} {[}\textit{arg2} {[}...{]}{]}{]} - run external
  \emph{plugin} (executable or function in shared library) with
  specified arguments. Execution of \emph{plugin} may not last longer
  than \emph{timeout} seconds.  If \emph{plugin} looks like
  \emph{function@path} then function \emph{int
    function(char{*},char{*},char{*},...)} from shared library
  \emph{path} is called (\emph{timeout} is not functional in that
  case). Rule matches if exit code is 0. In arguments following
  substitions are applied before plugin is started:

\begin{itemize}
\item \%D - subject of users's cerificate,
\item \%P - name of credentials' proxy file.
\end{itemize}

\item \textbf{\textit{lcas}}\textit{=library directory database} -
  call LCAS functions to check rule. Here \textit{library} is path to
  shared library of LCAS, either absolute or relative to
  \textit{directory}; \textit{directory} is path to LCAS installation
  directory, equivalent of LCAS\_DIR variable; \textit{database} is
  path to LCAS database, equivalent to LCAS\_DB\_FILE variable. Each
  arguments except \textit{library} is optional and may be either
  skiped or replaced with '{*}'.
\item \textbf{\textit{all}} - accept any user
\end{itemize}

@@

\begin{itemize}
\item If the identity is ``/O=Grid/OU=Wrong Place/CN=Bad Person'' it
  is rejected
\item If the identity is in the mapfile
  ``/etc/grid-security/internal-staff'' the identity is accepted
\item If the identity is a VOMS proxy with a nordugrid VO extension
  and admin capability, the identity is accepted
\end{itemize}

@@

The examples presented below contain a full configuration examples for
the GridFTP server. The A-REX and information system configurations
are not shown - this is described in other documents.

@@

A-REX serves as front-end to a batch-system that provides a
"low\_prio\_queue" and a "high\_prio\_queue". Assignment to the
different queues is done via local user identites.  More precisely,
the local users "smscg001, smscg002, smscg003" will be assigned to the
low\_prio\_queue, whereas users "smscgP001, smscgP002, smscgP003" to
the high\_prio\_queue (the configration of the batch-system to support
this is out of scope of this example).

Users sending jobs to A-REX should be assigned to one of the queues
depending on the credentials they present in their proxy
certificate. The assigment shall look as follows:
\begin{itemize}
\item /smscg , /smscg/Role=test , /smscg/Role=production => shall map
  to one of the smscg00[1-3] local identities (thus low\_prio\_queue)
\item /smscg/atlas , /smscg/atlas/Role=test ,
  /smscg/atlas/Role=production => shall map to one of the
  smscgP00[1-3] local identities (thus high\_prio\_queue)
\end{itemize}

@@

Notes:
\begin{itemize}
\item a DN based grid-mapfile is generated on the front-end with a
  default mapping entry for John. The grid-mapfile is only used by the
  information system (GIIS) to make the grid resource look eligible
  for jobs submitted by John.
\item the DN based grid-mapfile per se does not permit John to access
  the grid resource under different local identies (e.g. once as
  smscg001 and later as smscgP001), since the first matching DN
  defines the local identity John is to be mapped to. This is not a
  flaw since NorduGrid has support for lcmaps, which allows a
  're-mapping' of a user.
\item the mapping of the FQAN to the local user identity
  (e.g. "/smscg" to local user "smscg001") shall be done with lcmaps
  (in detail the lcmaps framework + lcmaps voms plugins). Direct VOMS
  based mapping is also possible.
\end{itemize}

@@

[group]
# In this authorization group we are going to check if user presents
# any proof that he belongs to 'smscg' VO. We can use that information 
# later to explicitely limit access to resources. If such access 
# control is not needed this group can be removed.
name="smscg_auth"

# Here we can use internal support of ARC for VOMS attrbutes
# voms="smscg * * *"
# If we want to limit access to resources also by other VOMS
# attributes then other voms rules similar to those defined
# below in [gridftpd] section may be used.

@@

# Or - probably prefered way in this use case - we can use 
# LCAS to analyze delegated proxy.
# First element after '=' sign is path to LCAS library whatever
# it is called in current implementation. Second is LCAS installation
# path - it will be used to set environment variable LCAS_DIR.
# And third element is path to LCAS database file - it will be passed
# to environemnt variable LCAS_DB_FILE.
# Function 'lcas_get_fabric_authorization' of specified LCAS library
# will be called with following 3 arguments
#  1. char* pointer to string containing DN of user
#  2. gss_cred_id_t variable pointing at delegated credentials of user
#  3. char* pointer to empty string
# Returned 0 int value is treated as positive response
lcas="/opt/glite/lib/liblcas.so /opt/glite /opt/glite/share/lcas.db"

@@

# By specifying 'no' here we limit users allowed to exatblish
# connection to this server to those specified in grid-mapfile. This
# may be not necessary if additional authorization is applied as done
# below. But this provides additional layer of protection so let it
# be.
allowunknown="no"

@@

# Or if we want to use all 6 local accounts and let mapping choose
# randomly within 2 group accounts 'simplepool' may be used. In
# example below 'unixgroup' ensures proper choise of group and
# 'simplepool' makes a choise from acoounts in pool. Last argument
# specifies directory containing file named 'pool'. That file contains
# list of local user accounts. Also this directory will be used for
# writing information about current mappings.
#unixgroup="smscg_high simplepool /var/nordugrid/smscg_high"
#unixgroup="smscg_low simplepool /var/nordugrid/smscg_low"

# And mapping prefered in this use case - through LCMAPS.  First
# element after '=' sign is path to LCMAPS library whatever it is
# called in current implementation. Second is LCMAPS installation path
# - it will be used to set environment variable LCMAPS_DIR.  And third
# element is path to LCMAPS database file - it will be passed to
# environemnt variable LCMAPS_DB_FILE. Those 3 arguments are followed
# list of policy names.
# Function 'lcmaps_run_and_return_username' of specified LCMAPS library
# will be called with following arguments
#  1. char* pointer to string containing DN of user
#  2. gss_cred_id_t variable pointing at delegated credentials of user
#  3. char* pointer to empty string
#  4. char** pointer for chosen username.
#  5. int variable containing number of policies
#  6. char** list of policy names
# Expected 0 int value returned and argument 4 set. Value returned in
# 4th argument is used as username of local account.
unixmap="* lcmaps /opt/glite/lib/liblcmaps.so /opt/glite \
  /opt/glite/share/lcmaps.db policy1 policy2"

# Here we can specify mapping to some harmless local user account for
# safety reasons. If that account is not allowed to submit jobs to
# LRMS then this will also work as authorization effectively cuting
# off users without proper VOMS attributes.
unixmap="nobody:nobody all"

>>>>

{\raggedleft NORDUGRID-TECH-26\par}

@@

With the evolution of Grid technologies, and the need for more
standardized interfaces to allow different Grid products to
communicate with each other, the Hosting Environment Daemon
(HED)~\cite{HED} framework and associated Advanced Resource Execution
Service (A-REX) web service were developed to replace the aging GFS
and GM with a modern standards-compliant product. However, the
transition to a new framework takes time and so the legacy GFS is
still supported and maintained until such time as it becomes
obsolete. This document is designed for administrators who wish to run
the GFS alongside A-REX and describes how to configure and run the GFS
service. For installation of the GFS, up to date instructions may be
found on the NorduGrid website \textbf{\url{http://www.nordugrid.org}}.

@@

To make GFS to inter-operate with other parts of ARC only one
\emph{jobplugin.so} needs to be configured.

@@

\begin{itemize}
\item \textbf{\textit{pidfile}}\textit{=path} -- specifies file where
  process id of the gridftpd process will be stored. Defaults to
  \emph{/var/run/gridftpd.pid} if running as root and
  \emph{\$HOME/gridftpd.pid} otherwise.
\item \textbf{\textit{logfile}}\textit{=path} -- specifies name of file
  for logging debug/informational output. Defaults to
  \emph{/var/log/gridftpd.log}.
\item \textbf{\textit{logsize}}\textit{=size number} -- restricts
  log file size to \emph{size} and keeps \emph{number} archived log
  files.
\item \textbf{\textit{debug}}\textit{=number} -- specifies level of
  debug information. More information is printed for higher
  levels. Currently the highest effective number is 5 (DEBUG) and
  lowest 0 (FATAL). Defaults to 3 (INFO).
\item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port
  number.  Default is 2811.
\item \textbf{\textit{include}}\textit{=path} - include contents of
  another file. Generic commands cannot be specified there.
\item \textbf{\textit{encryption}}\textit{=yes|no} - specifies if
  server will allow data transfer to be encrypted. Default is yes.
\item \textbf{\textit{pluginpath}}\textit{=path} - specifies the path
  where plugin libraries are installed. In a normal installation this
  is \emph{\$(ARC\_LOCATION)/lib/arc}.
\item \textbf{\textit{allowunknown}}\textit{=yes|no} - if set to
  \emph{yes}, clients are not checked against the grid-mapfile. Hence
  only access rules specified in this configuration file will be
  applied.
\item \textbf{\textit{firewall}}\textit{=hostname} - use IP address of
  the \textit{hostname} in response to PASV command instead of IP
  address of a network interface of the computer. An IP address can be
  used instead of \textit{hostname}. This command may be useful if the
  server is situated behind a NAT.
\item \textbf{\textit{unixgroup}}\textit{=group rule} - define local
  UNIX user and optionally UNIX group to which user belonging to
  specified authorization \textit{group} is mapped (see
  Section~\ref{sec:Authorization} for definition of group). Local
  names are obtained from the specified \textit{rule}. If the
  specified rule could not produce any mapping, the next command is
  used. Mapping stops at first matched rule. The following rules are
  supported:

@@

\begin{itemize}
\item \%D - subject of user's certificate,
\item \%P - name of credentials' proxy file.
\end{itemize}

@@

\begin{itemize}
\item {[}\textbf{\textit{subject}}{]}\textit{=subject}
  {[}\textit{subject} {[}...{]}{]} - match user with one of specified
  subjects
\item \textbf{\textit{file}}={[}\textit{filename} {[}...{]}{]} - read
  rules from specified files (format of file is similar to Globus
  grid-mapfile with user names ignored)
\item \textbf{\textit{remote}}={[}\textit{ldap://host:port/dn}
  {[}...{]}{]} - match user listed in one of specified LDAP
  directories (uses network connection hence can take time to process)
\item \textbf{\textit{voms}}\textit{=vo group role capabilities} -
  accept user with VOMS proxy with specified \textit{vo},
  \textit{group}, \textit{role} and \textit{capabilities}. '{*}' can
  be used to accept any value
\item \textbf{\textit{vo}}={[}\textit{vo} {[}...{]}{]} - match user
  belonging to one of specified Virtual Organizations as defined in
  \emph{vo} configuration section (see below).
\item \textbf{\textit{group}}={[}\textit{groupname}
  {[}\textit{groupname} {[}...{]}{]}{]} - match user already belonging
  to one of specified groups.
\item \textbf{\textit{plugin}}\textit{=timeout plugin}
  {[}\textit{arg1} {[}\textit{arg2} {[}...{]}{]}{]} - run external
  \emph{plugin} (executable or function in shared library) with
  specified arguments. Execution of \emph{plugin} may not last longer
  than \emph{timeout} seconds.  If \emph{plugin} looks like
  \emph{function@path} then function \emph{int
    function(char{*},char{*},char{*},...)} from shared library
  \emph{path} is called (\emph{timeout} is not functional in that
  case). Rule matches if exit code is 0. In arguments following
  substitutions are applied before plugin is started:

\begin{itemize}
\item \%D - subject of user's certificate,
\item \%P - name of credentials' proxy file.
\end{itemize}

\item \textbf{\textit{lcas}}\textit{=library directory database} -
  call LCAS functions to check rule. Here \textit{library} is path to
  shared library of LCAS, either absolute or relative to
  \textit{directory}; \textit{directory} is path to LCAS installation
  directory, equivalent of LCAS\_DIR variable; \textit{database} is
  path to LCAS database, equivalent to LCAS\_DB\_FILE variable. Each
  arguments except \textit{library} is optional and may be either
  skipped or replaced with '{*}'.
\item \textbf{\textit{all}} - accept any user
\end{itemize}

@@

\begin{itemize}
\item If the identity is ``/O=Grid/OU=Wrong Place/CN=Bad Person'' it
  is rejected
\item If the identity is in the mapfile
  ``/etc/grid-security/internal-staff'' the identity is accepted
\item If the identity is a VOMS proxy with a NorduGrid VO extension
  and admin capability, the identity is accepted
\end{itemize}

@@

The examples presented below contain full configuration examples for
the GridFTP server. The A-REX and information system configurations
are not shown - this is described in other documents.

@@

A-REX serves as front-end to a batch-system that provides a
"low\_prio\_queue" and a "high\_prio\_queue". Assignment to the
different queues is done via local user identites.  More precisely,
the local users "smscg001, smscg002, smscg003" will be assigned to the
low\_prio\_queue, whereas users "smscgP001, smscgP002, smscgP003" to
the high\_prio\_queue (the configuration of the batch-system to support
this is out of scope of this example).

Users sending jobs to A-REX should be assigned to one of the queues
depending on the credentials they present in their proxy
certificate. The assignment shall look as follows:
\begin{itemize}
\item /smscg , /smscg/Role=test , /smscg/Role=production => shall map
  to one of the smscg00[1-3] local identities (thus low\_prio\_queue)
\item /smscg/atlas , /smscg/atlas/Role=test ,
  /smscg/atlas/Role=production => shall map to one of the
  smscgP00[1-3] local identities (thus high\_prio\_queue)
\end{itemize}

@@

Notes:
\begin{itemize}
\item a DN based grid-mapfile is generated on the front-end with a
  default mapping entry for John. The grid-mapfile is only used by the
  information system (GIIS) to make the grid resource look eligible
  for jobs submitted by John.
\item the DN based grid-mapfile per se does not permit John to access
  the grid resource under different local identities (e.g. once as
  smscg001 and later as smscgP001), since the first matching DN
  defines the local identity John is to be mapped to. This is not a
  flaw since NorduGrid has support for lcmaps, which allows a
  're-mapping' of a user.
\item the mapping of the FQAN to the local user identity
  (e.g. "/smscg" to local user "smscg001") shall be done with lcmaps
  (in detail the lcmaps framework + lcmaps voms plugins). Direct VOMS
  based mapping is also possible.
\end{itemize}

@@

[group]
# In this authorization group we are going to check if user presents
# any proof that he belongs to 'smscg' VO. We can use that information 
# later to explicitly limit access to resources. If such access 
# control is not needed this group can be removed.
name="smscg_auth"

# Here we can use internal support of ARC for VOMS attributes
# voms="smscg * * *"
# If we want to limit access to resources also by other VOMS
# attributes then other voms rules similar to those defined
# below in [gridftpd] section may be used.

@@

# Or - probably preferred way in this use case - we can use 
# LCAS to analyze delegated proxy.
# First element after '=' sign is path to LCAS library whatever
# it is called in current implementation. Second is LCAS installation
# path - it will be used to set environment variable LCAS_DIR.
# And third element is path to LCAS database file - it will be passed
# to environment variable LCAS_DB_FILE.
# Function 'lcas_get_fabric_authorization' of specified LCAS library
# will be called with following 3 arguments
#  1. char* pointer to string containing DN of user
#  2. gss_cred_id_t variable pointing at delegated credentials of user
#  3. char* pointer to empty string
# Returned 0 int value is treated as positive response
lcas="/opt/glite/lib/liblcas.so /opt/glite /opt/glite/share/lcas.db"

@@

# By specifying 'no' here we limit users allowed to establish
# connection to this server to those specified in grid-mapfile. This
# may be not necessary if additional authorization is applied as done
# below. But this provides additional layer of protection so let it
# be.
allowunknown="no"

@@

# Or if we want to use all 6 local accounts and let mapping choose
# randomly within 2 group accounts 'simplepool' may be used. In
# example below 'unixgroup' ensures proper choice of group and
# 'simplepool' makes a choise from accounts in pool. Last argument
# specifies directory containing file named 'pool'. That file contains
# list of local user accounts. Also this directory will be used for
# writing information about current mappings.
#unixgroup="smscg_high simplepool /var/nordugrid/smscg_high"
#unixgroup="smscg_low simplepool /var/nordugrid/smscg_low"

# And mapping preferred in this use case - through LCMAPS.  First
# element after '=' sign is path to LCMAPS library whatever it is
# called in current implementation. Second is LCMAPS installation path
# - it will be used to set environment variable LCMAPS_DIR.  And third
# element is path to LCMAPS database file - it will be passed to
# environment variable LCMAPS_DB_FILE. Those 3 arguments are followed
# list of policy names.
# Function 'lcmaps_run_and_return_username' of specified LCMAPS library
# will be called with following arguments
#  1. char* pointer to string containing DN of user
#  2. gss_cred_id_t variable pointing at delegated credentials of user
#  3. char* pointer to empty string
#  4. char** pointer for chosen username.
#  5. int variable containing number of policies
#  6. char** list of policy names
# Expected 0 int value returned and argument 4 set. Value returned in
# 4th argument is used as username of local account.
unixmap="* lcmaps /opt/glite/lib/liblcmaps.so /opt/glite \
  /opt/glite/share/lcmaps.db policy1 policy2"

# Here we can specify mapping to some harmless local user account for
# safety reasons. If that account is not allowed to submit jobs to
# LRMS then this will also work as authorization effectively cutting
# off users without proper VOMS attributes.
unixmap="nobody:nobody all"
